---
title: "Compilation of data for downstream use"
author: "Daniel Skubleny"
date: '2023-01-05'
output: html_document
---


#Import libraries
```{r, warning= FALSE, message=FALSE}
#Libraries
library(readxl)
library(ggplot2)
library(devtools)
library(Biobase)
library(AnnotationDbi)
library(BiocManager)
library(plyr)
library(dplyr)
library(annotate)
library(stringr)
library(org.Hs.eg.db)
library(GEOquery)
library(readr)
library(survival)
library(survminer)
library(data.table)
library(RColorBrewer)
library(Rmisc)
library(FSA)
library(openxlsx)
library(maftools)
library(TCGAbiolinks)
```

#Import and arrange data
```{r}
#Import data
## Import Data

immune_landscape <- read.csv("Immune Landscape of Cancer/1-s2.0-S1074761318301213-mmc2.xlsx")
patient_data <- read.csv("nmf_coefficient_rank100_PanCancer_10274cases.csv")
```
```{r}
#Clean and arrange data
immune_df = dplyr::select(immune_landscape, c("TCGA Participant Barcode", "Immune Subtype", "TCGA Subtype", "Aneuploidy Score"))

# Change column names 
colnames(immune_df)[which(names(immune_df) == "TCGA Participant Barcode")] <- "Patient_ID"
immune_df[,1:ncol(immune_df)] = lapply(1:ncol(immune_df),function(x) as.factor(immune_df[[x]]))
immune_df = na_if(immune_df, 'NA')
immune_df <- immune_df %>% filter("Immune Subtype" != 'NA') %>% droplevels()
```

#Load and prepare expression data 
```{r}
edata <- read.delim("EBPlusPlusAdjustPANCAN_IlluminaHiSeq_RNASeqV2.geneExp.tsv")

sampleid <- read.csv("nmf_coefficient_rank100_PanCancer_10274cases.csv")
```

#GENE LEVEL DATA
```{r}
NMF_genes <- read.csv("nmf_basis_rank100_PanCancer.csv", row.names=1)
```

```{r}
genenames = NMF_genes$Gene_Id
edata = edata %>% tibble::column_to_rownames("gene_id") # Make the first column the row name
edata <- subset(edata, rownames(edata) %in% as.matrix(genenames))

#Visualize data = boxplot
par(mfrow=c(1,2))
boxplot(edata[,1:50], col=2, range=0)
boxplot(log2(edata+1)[,1:50], col=2, range=0)
#Visualize data = barplot
par(mfrow=c(1,2))
hist(edata[,1],col=2)
hist(log2(edata[,1]+1),col=3)
# Bland Altman Plots
mm = edata[,1] - edata[,2]
aa = edata[,1] + edata[,2]
plot(aa,mm,col=2) 

mm = log2(edata[,1]+1) - log2(edata[,2]+1)
aa = log2(edata[,1]+1) + log2(edata[,2]+1)
plot(aa,mm,col=2)
par(mfrow=c(1,1))

#These visualizations confirm that log2+1 transformation is acceptable.

tedata = t(edata)
tedata = as.data.frame(tedata)
```
```{r}
#Select Sample ID and replace "-" with "."
sampleid = dplyr::select(sampleid, c("Sample_ID","Patient_ID"))

df = sampleid$Sample_ID = str_replace_all(sampleid$Sample_ID, c("-" = "."))
tedata <- subset(tedata, rownames(tedata) %in% as.matrix(df))
tedata = tedata %>% tibble::rownames_to_column("Sample_ID") # Make rownames the first column

#Modify matrix orientation
tedata = merge(sampleid,tedata, by.x="Sample_ID")
tedata = dplyr::select(tedata, -"Sample_ID") # Remove Sample_ID columns
```
```{r}
#Log2 transform expression data
exprs_data = tedata
exprs_data[,2:16336] = log2(exprs_data[,2:16336] +1)
```

#Clinical Data
```{r}
tcga_clinical <- read_excel("TCGA-CDR-SupplementalTableS1.xlsx")

tcga_clinical = dplyr::select(tcga_clinical, c("bcr_patient_barcode","type","age_at_initial_pathologic_diagnosis","gender","race","ajcc_pathologic_tumor_stage","clinical_stage", "histological_grade","treatment_outcome_first_course", "new_tumor_event_type", "DFI"))

colnames(tcga_clinical)[which(names(tcga_clinical) == "bcr_patient_barcode")] <- "Patient_ID"
```

#Load all data for TCGA studies included in the survival analysis. LAML has entirely different data and specific therapy techniques--- will exclude
####BRCA
```{r}
query <- GDCquery(project = "TCGA-BRCA", 
                  data.category = "Clinical",
                  data.type = "Clinical Supplement", 
                  data.format = "BCR Biotab")
GDCdownload(query)
brca.clinical.BCRtab.all <- GDCprepare(query)
```
#Breast Cancer data 
```{r}
brca.clinical.data = brca.clinical.BCRtab.all[["clinical_patient_brca"]]
brca.clinical.data = brca.clinical.data[-c(1:2),]
brca.clinical.data = dplyr::select(brca.clinical.data, c("bcr_patient_barcode","her2_ihc_score","pr_status_by_ihc", "er_status_by_ihc", "her2_fish_status"))
colnames(brca.clinical.data)[which(names(brca.clinical.data) == "bcr_patient_barcode")] <- "Patient_ID"
```
#TME Score
```{r}
TMEscore <- read_excel("tmescore_withstad.xlsx", sheet = "S13", col_names = FALSE, col_types = c("text", "text", "text", "text", "numeric", "numeric", "numeric", "text"))
```
```{r}
#Set column names 
colnames(TMEscore) = NULL
colnames(TMEscore) = TMEscore[2,]
TMEscore = TMEscore[-c(1:2),]

colnames(TMEscore)[which(names(TMEscore) == "ID")] <- "Patient_ID"
```

#Tumor mutational burden
```{r}
#Using TMB is a proxy for two potential mechanisms. One is the relationship of TMB to the efficacy of DNA repair mechanisms. Factors associated with higher TMB would be definicient in DNA repair mechanisms. The second is the relationship of TMB to immunotherapy response. FDA has approved pembrolizumab for TMB>10 according to a targeted assay. Here we are using WES data and apply the same cut-off value. 
#Phenotype Data
```

```{r}
laml.maf <- system.file("extdata", "tcga_laml.maf.gz", package = "maftools")
laml <- read.maf(maf = laml.maf)

tmb = tcgaCompare(maf=laml,logscale = TRUE)
tmb_sample = tmb[["mutation_burden_perSample"]]
```
```{r}
#Fix patient ID
tmb_sample$Tumor_Sample_Barcode = as.character(tmb_sample$Tumor_Sample_Barcode)
Patient_ID = substr(tmb_sample$Tumor_Sample_Barcode,1,nchar(tmb_sample$Tumor_Sample_Barcode)-16)
tmb_sample$Patient_ID = Patient_ID
tmb_sample$Patient_ID = as.factor(tmb_sample$Patient_ID)
tmb_sample = dplyr::select(tmb_sample, -"Tumor_Sample_Barcode")
tmb_sample$logtmb = log10(tmb_sample$total)
```




```{r}
pheno_data = patient_data
pheno_data = dplyr::select(pheno_data, "Patient_ID", "Sample_ID")
pheno_data = merge(pheno_data,tcga_clinical, by.x="Patient_ID", all = TRUE)
pheno_data = pheno_data[!is.na(pheno_data$Sample_ID),]
pheno_data = merge(pheno_data,brca.clinical.data, by.x="Patient_ID", all = TRUE)
pheno_data = pheno_data[!is.na(pheno_data$Sample_ID),]
pheno_data = merge(pheno_data,immune_df, by.x="Patient_ID", all = TRUE)
pheno_data = pheno_data[!is.na(pheno_data$Sample_ID),]
tme.score = dplyr::select(TMEscore, c("Patient_ID","TMEscore_binary"))
pheno_data = merge(pheno_data,tme.score, by.x="Patient_ID", all = TRUE)
pheno_data = pheno_data[!is.na(pheno_data$Sample_ID),]
mutation_burden = dplyr::select(tmb_sample, c("Patient_ID", "total"))
#Aggregate duplicate TMB by mean
mutation_burden %>% group_by(Patient_ID) %>% summarise_all(mean) %>% data.frame() -> mutation_burden
pheno_data = merge(pheno_data,mutation_burden, by.x="Patient_ID", all = TRUE)
pheno_data = pheno_data[!is.na(pheno_data$Sample_ID),]
```
```{r}
#Make missing data NA, remove data with limited use.
pheno_data = na_if(pheno_data, '[Unknown]')
pheno_data = na_if(pheno_data, '[Not Available]')
pheno_data = na_if(pheno_data, '[Not Available]|[Not Available]')
pheno_data = na_if(pheno_data, '[Not Applicable]')
pheno_data = na_if(pheno_data, '')
pheno_data = na_if(pheno_data, '[Discrepancy]')
pheno_data = na_if(pheno_data, '[Not Evaluated]')
pheno_data = na_if(pheno_data, "Stage X")
pheno_data = na_if(pheno_data, "I or II NOS")
pheno_data = na_if(pheno_data, "Stage 0")
pheno_data = na_if(pheno_data, "Stage IS")
pheno_data = na_if(pheno_data, "IS")
pheno_data = na_if(pheno_data, "I/II NOS")
pheno_data = na_if(pheno_data, "NOS")

#Make new variable stage. Prefer pathologic stage if present. If NA use clinical stage if present
pheno_data = pheno_data %>% 
  mutate(Stage = if_else(is.na(ajcc_pathologic_tumor_stage), clinical_stage,ajcc_pathologic_tumor_stage))

#Code new_tumor_event_type as yes and no. Exclude new primary. 
#Split histology into Lauren class and WHO class
pheno_data$New_tumour_event <- ifelse(grepl("Distant", pheno_data$new_tumor_event_type, ignore.case = T), "Yes",
                              ifelse(grepl("Locoregional", pheno_data$new_tumor_event_type, ignore.case = T), "Yes",
                              ifelse(grepl("Recurrence", pheno_data$new_tumor_event_type, ignore.case = T), "Yes",
                              ifelse(grepl("Progression", pheno_data$new_tumor_event_type, ignore.case = T), "Yes",
                              ifelse(grepl("Metastatic", pheno_data$new_tumor_event_type, ignore.case = T), "Yes",
                              ifelse(grepl("Regional", pheno_data$new_tumor_event_type, ignore.case = T),"Yes",
                              ifelse(grepl("Biochemical", pheno_data$new_tumor_event_type, ignore.case = T), "Yes",
                              ifelse(grepl("New primary", pheno_data$new_tumor_event_type, ignore.case = T), "Yes", "No"))))))))

#Recurrence
pheno_data$DFI <- ifelse(pheno_data$DFI == 1, "Yes", "No")

#Pathologic Stage
pheno_data <- pheno_data %>% filter("Stage" != '') %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != '[Discrepancy]') %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != '[Unknown]') %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != '[Not Available]') %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != '[Discrepancy]') %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != "Stage X") %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != "I or II NOS") %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != "Stage 0") %>% droplevels()
pheno_data <- pheno_data %>% filter("Stage" != "Stage IS") %>% droplevels()

#Convert strings
pheno_data$Stage<- ifelse(grepl("IV", pheno_data$Stage, ignore.case = T), "Stage IV",
                              ifelse(grepl("III", pheno_data$Stage, ignore.case = T), "Stage III",
                              ifelse(grepl("II", pheno_data$Stage, ignore.case = T), "Stage II", 
                              ifelse(grepl("I", pheno_data$Stage, ignore.case = T), "Stage I", "No Stage"))))
pheno_data = na_if(pheno_data, 'No Stage')

#Race
pheno_data <- pheno_data %>% filter("race" != '[Unknown]') %>% droplevels()
pheno_data <- pheno_data %>% filter("race" != '[Not Available]') %>% droplevels()
pheno_data <- pheno_data %>% filter("race" != '[Not Evaluated]') %>% droplevels()

#Grade
pheno_data = na_if(pheno_data, "GB") 
pheno_data = na_if(pheno_data, "High Grade") 
pheno_data = na_if(pheno_data, "Low Grade") 
pheno_data = na_if(pheno_data, "GX") 

pheno_data <- pheno_data %>% filter("histological_grade" != '') %>% droplevels()
pheno_data <- pheno_data %>% filter("histological_grade" != '[Discrepancy]') %>% droplevels()
pheno_data <- pheno_data %>% filter("histological_grade" != '[Unknown]') %>% droplevels()
pheno_data <- pheno_data %>% filter("histological_grade" != '[Not Available]') %>% droplevels()
pheno_data <- pheno_data %>% filter("histological_grade" != 'GB') %>% droplevels()
pheno_data <- pheno_data %>% filter("histological_grade" != 'High Grade') %>% droplevels()
pheno_data <- pheno_data %>% filter("histological_grade" != 'Low Grade') %>% droplevels()

#Breast CA
pheno_data <- pheno_data %>% filter("pr_status_by_ihc" != '[Not Evaluated]') %>% droplevels()
pheno_data <- pheno_data %>% filter("her2_ihc_score" != '[Not Available]') %>% droplevels()
pheno_data <- pheno_data %>% filter("er_status_by_ihc" != '[Not Evaluated]') %>% droplevels()
pheno_data <- pheno_data %>% filter("her2_fish_status" != '[Not Evaluated]') %>% droplevels()
pheno_data <- pheno_data %>% filter("her2_fish_status" != '[Not Available]') %>% droplevels()
```
```{r}
#Change column names 
colnames(pheno_data)[which(names(pheno_data) == "type")] <- "Type"
colnames(pheno_data)[which(names(pheno_data) == "gender")] <- "Sex"
colnames(pheno_data)[which(names(pheno_data) == "race")] <- "Race"
colnames(pheno_data)[which(names(pheno_data) == "age_at_initial_pathologic_diagnosis")] <- "Age"
colnames(pheno_data)[which(names(pheno_data) == "histological_grade")] <- "Grade"
colnames(pheno_data)[which(names(pheno_data) == "TMEscore_binary")] <- "Tumour_Microenvironment_Score"
colnames(pheno_data)[which(names(pheno_data) == "total")] <- "Tumour_Mutational_Burden"
colnames(pheno_data)[which(names(pheno_data) == "DFI")] <- "Recurrence"
colnames(pheno_data)[which(names(pheno_data) == "TCGA Subtype")] <- "TCGA_Subtype"
colnames(pheno_data)[which(names(pheno_data) == "Aneuploidy Score")] <- "Aneuploidy_Score"
colnames(pheno_data)[which(names(pheno_data) == "Immune Subtype")] <- "Immune_Subtype"

#Remove useless columns
pheno_data = dplyr::select(pheno_data, -c("ajcc_pathologic_tumor_stage","clinical_stage","new_tumor_event_type"))
```

```{r}
pheno_data$Stage = as.factor(pheno_data$Stage)
pheno_data$Type = as.factor(pheno_data$Type)
pheno_data$Sex = as.factor(pheno_data$Sex)
pheno_data$Race = as.factor(pheno_data$Race)
pheno_data$Grade = as.factor(pheno_data$Grade)
pheno_data$Recurrence = as.factor(pheno_data$Recurrence)
pheno_data$her2_ihc_score = as.factor(pheno_data$her2_ihc_score)
pheno_data$pr_status_by_ihc = as.factor(pheno_data$pr_status_by_ihc)
pheno_data$er_status_by_ihc = as.factor(pheno_data$er_status_by_ihc)
pheno_data$her2_fish_status = as.factor(pheno_data$her2_fish_status)
pheno_data$Tumour_Microenvironment_Score = as.factor(pheno_data$Tumour_Microenvironment_Score)
pheno_data$New_tumour_event = as.factor(pheno_data$New_tumour_event)
```

#Survival Data subset

```{r}
# Survival data ----------------------------------------------------------------
surv_data <- fread(
  file = "survivaldata_7922cases.csv",
  sep = ",",
  header = T,
  na.strings = NULL,
  stringsAsFactors = F,
  logical01 = T,
  data.table = F
)
row.names(surv_data) <- surv_data$Sample_ID
surv_data$Sample_ID <- NULL
```
```{r}
survival_data = surv_data 
survival_data$vital_status = as.factor(survival_data$vital_status)
survival_data$vital_status <- ifelse(survival_data$vital_status == "Dead", 2, 1)

survival_data = dplyr::select(survival_data, c("Patient_ID", "vital_status", "OS_time", "OS_censor"))
survival_data = merge(survival_data,pheno_data, by.x="Patient_ID")
survival_data = survival_data[!is.na(survival_data$Sample_ID),]

survival_data<-survival_data[!(survival_data$OS_censor==-9),]

```

```{r}
write.csv(pheno_data,'NMF_clinical_data_final.csv')
write.csv(survival_data,'NMF_survival_data_final.csv')
```
#Combine pheno data and NMF factors values for models
```{r}
factors = patient_data
factors = dplyr::select(factors, -"Sample_ID")
#Phenotype combo with all 10274 patients
pheno_combo = merge(factors,pheno_data, by.x="Patient_ID")
#Survival combo with only the 7922 survival data patients
survival_combo = merge(factors,survival_data, by.x="Patient_ID")
```

pheno_data and survival_data are the main data frames used for biological analysis.

They are saved in the repo as NMF_clinical_data_final.csv and NMF_survival_data_final.csv, respectively. 




