---
title: "NMF Factor Random Forest Models"
author: "Daniel Skubleny"
date: '2023-01-07'
output: html_document
---

#Import libraries
```{r, warning= FALSE, message=FALSE}
#Libraries
library(readxl)
library(ggplot2)
library(devtools)
library(Biobase)
library(AnnotationDbi)
library(BiocManager)
library(plyr)
library(dplyr)
library(sva)
library(snpStats)
library(limma)
library(edge)
library(annotate)
library(sqldf)
library(stringr)
library(randomForest)
library(caret)
library(preprocessCore)
library(glmnet)
library(klaR)
library(gbm)
library(org.Hs.eg.db)
library(measures)
library(GEOquery)
library(remotes)
library(naivebayes)
library(mlr)
library(mlr3filters)
library(Metrics)
library(MLmetrics)
library(readr)
library(pheatmap)
library(survival)
library(survminer)
library(ggRandomForests)
library(data.table)
library(fmsb)
library(RColorBrewer)
library(Rmisc)
library(FSA)
library(openxlsx)
library(maftools)
library(MuMIn)
library(ggVennDiagram)
library(grid)
library(ggeffects)
library(gratia)
```

#Load survival data as survival_data
```{r}
survival_data <- read.csv("NMF_survival_data_final.csv", row.names=1)

#Combine the dataframe to incorporate 100 NMF Factors 
patient_data <- read.csv("nmf_coefficient_rank100_PanCancer_10274cases.csv")

factors = patient_data
factors = dplyr::select(factors, -"Sample_ID")
#Survival combo with only the 7922 survival data patients
survival_combo = merge(factors,survival_data, by.x="Patient_ID")
```

#Prop test for ACC>NIR analysis 
```{r}
propTest <- function(x){
    res <- try(
      binom.test(sum(diag(x)),
                 sum(x),
                 p = max(apply(x, 2, sum)/sum(x)),
                 alternative = "greater"),
      silent = TRUE)
    res <- if(inherits(res, "try-error"))
      c("null.value.probability of success" = NA, p.value = NA)
    else
      res <- unlist(res[c("null.value", "p.value")])
    res
}
```


#Model Immune subtypes
####Survival Cases
```{r}
#Model NMF representation by immune subtypes. 
immune_merge_survival = survival_combo[!is.na(survival_combo$Immune_Subtype),]

#Make our test and train data
x= data.matrix(immune_merge_survival[,2:101])
y= immune_merge_survival$Immune_Subtype
y = as.factor(y)
```
```{r}
set.seed(99)
rf.NMF.immune_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
importance = varImp(rf.NMF.immune_survival)
```
```{r}
immune_info = as.data.frame(table(immune_merge_survival$Immune_Subtype))
sum(immune_info$Freq)
length(immune_info$Var1)
summary(immune_info$Freq)
```
```{r}
immune_confmat = confusionMatrix(rf.NMF.immune_survival, norm="none")
immune_proptest = propTest(immune_confmat$table)
immune_proptest
```
#TCGA subtypes
####Survival Cases
```{r}
#Model NMF representation by all TCGA subtypes subtypes. 
subtypes_merge_survival = survival_combo[!is.na(survival_combo$TCGA_Subtype),]
subtypes_merge_survival$TCGA_Subtype = factor(subtypes_merge_survival$TCGA_Subtype)

#Make our test and train data
x= data.matrix(subtypes_merge_survival[,2:101])
y= subtypes_merge_survival$TCGA_Subtype
y = as.factor(y)
```
```{r}
set.seed(99)
rf.NMF.subtype_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5,savePredictions="all"))
```
```{r}
importance = varImp(rf.NMF.subtype_survival)
```
```{r}
subtypes_info = as.data.frame(table(subtypes_merge_survival$TCGA_Subtype))
sum(subtypes_info$Freq)
length(subtypes_info$Var1)
summary(subtypes_info$Freq)
```
```{r}
subtypes_confmat = confusionMatrix(rf.NMF.subtype_survival, norm="none")
subtypes_proptest = propTest(subtypes_confmat$table)
```

###CHECK CODE BELOW
#TME Score evaluation 
####Survival Cases
```{r}
#Model NMF representation by aneuploidy score 
TMEscore_merge_survival = survival_combo[!is.na(survival_combo$Tumour_Microenvironment_Score),]
TMEscore_merge_survival$Tumour_Microenvironment_Score = factor(TMEscore_merge_survival$Tumour_Microenvironment_Score)

#Make our test and train data
x= data.matrix(TMEscore_merge_survival[,2:101])
y= TMEscore_merge_survival$Tumour_Microenvironment_Score
y = as.factor(y)
#Relevel for that High TME score is coded as 1
y <- relevel(y, "Low")
```
```{r}
set.seed(99)
rf.TME_all_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
importance = varImp(rf.TME_all_survival)
```
```{r}
TME_info = as.data.frame(table(TMEscore_merge_survival$Tumour_Microenvironment_Score))
sum(TME_info$Freq)
length(TME_info$Var1)
summary(TME_info$Freq)
```
```{r}
TME_confmat = confusionMatrix(rf.TME_all_survival, norm="none")
TME_proptest = propTest(TME_confmat$table)
TME_proptest
```


#Organ merge
####Survival
```{r}
#Model NMF representation by organ of origin. 
organ_merge_survival = survival_combo[!is.na(survival_combo$Type),]
organ_merge_survival$Type = factor(organ_merge_survival$Type)

#Make our test and train data
x= data.matrix(organ_merge_survival[,2:101])
y= organ_merge_survival$Type
y = as.factor(y)
```
```{r}
set.seed(99)
rf.NMF.organ_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
importance.organ = varImp(rf.NMF.organ_survival)
```
```{r}
organ_info = as.data.frame(table(organ_merge_survival$Type))
sum(organ_info$Freq)
length(organ_info$Var1)
summary(organ_info$Freq)
```
```{r}
organ_confmat = confusionMatrix(rf.NMF.organ_survival, norm="none")
organ_proptest = propTest(organ_confmat$table)
organ_confmat
organ_proptest
```

#Model Pathologic stage, trimmed
####Survival cases
```{r}
#Model NMF representation by Stage
stage_survival = survival_combo[!is.na(survival_combo$Stage),]
stage_survival$Stage = factor(stage_survival$Stage)

#Make our test and train data
x= data.matrix(stage_survival[,2:101])
y= stage_survival$Stage
y = as.factor(y)
```
```{r}
set.seed(99)
rf.stage_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
importance = varImp(rf.stage_survival)
```
```{r}
stage_info = as.data.frame(table(stage_survival$Stage))
sum(stage_info$Freq)
length(stage_info$Var1)
summary(stage_info$Freq)
```
```{r}
stage_confmat = confusionMatrix(rf.stage_survival, norm="none")
stage_proptest = propTest(stage_confmat$table)
stage_confmat
stage_proptest
```
#Model Histologic grade
####Survival cases
```{r}
#Model NMF representation by histologic grade
grade_survival = survival_combo[!is.na(survival_combo$Grade),]
grade_survival$Grade = factor(grade_survival$Grade)

#Make our test and train data
x= data.matrix(grade_survival[,2:101])
y= grade_survival$Grade
y = as.factor(y)
```
```{r}
set.seed(99)
rf.grade_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
grade_info = as.data.frame(table(grade_survival$Grade))
sum(grade_info$Freq)
length(grade_info$Var1)
summary(grade_info$Freq)
```
```{r}
grade_confmat = confusionMatrix(rf.grade_survival, norm="none")
grade_proptest = propTest(grade_confmat$table)
grade_confmat
grade_proptest
```
#Model sex
####Survival cases
```{r}
#Model NMF representation by sex
sex_survival = survival_combo[!is.na(survival_combo$Sex),]
sex_survival$Sex = factor(sex_survival$Sex)

#Make our test and train data
x= data.matrix(sex_survival[,2:101])
y= sex_survival$Sex
y = as.factor(y)
```
```{r}
set.seed(99)
rf.sex_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
sex_info = as.data.frame(table(sex_survival$Sex))
sum(sex_info$Freq)
length(sex_info$Var1)
summary(sex_info$Freq)
```
```{r}
sex_confmat = confusionMatrix(rf.sex_survival, norm="none")
sex_proptest = propTest(sex_confmat$table)
sex_confmat
sex_proptest
```
#Model Age
####Survival cases
```{r}
#Merge expression data and subtype allocation
clinical_age_merge_survival = survival_combo
#Split age into relevant factor ranges
clinical_age_merge_survival$Age <- cut(clinical_age_merge_survival$Age, c(0,40,Inf), c("Age < 40","Age >40"))
clinical_age_merge_survival = clinical_age_merge_survival[!is.na(clinical_age_merge_survival$Age),]

```
```{r}
#Model NMF representation by immune subtypes. 

#Make our test and train data
x= data.matrix(clinical_age_merge_survival[,2:101])
y= clinical_age_merge_survival$Age
y = as.factor(y)
```
```{r}
set.seed(99)
rf.age_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
age_info = as.data.frame(table(clinical_age_merge_survival$Age))
sum(age_info$Freq)
length(age_info$Var1)
summary(age_info$Freq)
```
```{r}
age_confmat = confusionMatrix(rf.age_survival, norm="none")
age_proptest = propTest(age_confmat$table)
age_confmat
age_proptest
```

#Model reccurent event
####Survival cases
```{r}
#Model NMF representation by recurrent event
recurrence_survival = survival_combo[!is.na(survival_combo$Recurrence),]
recurrence_survival$Recurrence = factor(recurrence_survival$Recurrence)

#Make our test and train data
x= data.matrix(recurrence_survival[,2:101])
y= recurrence_survival$Recurrence
y = as.factor(y)
```
```{r}
set.seed(99)
rf.recurrent_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
recurrence_info = as.data.frame(table(recurrence_survival$Recurrence))
sum(recurrence_info$Freq)
length(recurrence_info$Var1)
summary(recurrence_info$Freq)
```
```{r}
recurrence_confmat = confusionMatrix(rf.recurrent_survival, norm="none")
recurrence_proptest = propTest(recurrence_confmat$table)
recurrence_confmat
recurrence_proptest
```

#Model New tumour event
####Survival cases
```{r}
#Model NMF representation by new tumour event
nte_survival = survival_combo[!is.na(survival_combo$New_tumour_event),]
nte_survival$New_tumour_event = factor(nte_survival$New_tumour_event)

#Make our test and train data
x= data.matrix(nte_survival[,2:101])
y= nte_survival$New_tumour_event
y = as.factor(y)
```
```{r}
set.seed(99)
rf.nte_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```

#Model race
####Survival cases
```{r}
#Model NMF representation by immune subtypes. 
race_survival = survival_combo[!is.na(survival_combo$Race),]
race_survival$Race = factor(race_survival$Race)

#Make our test and train data
x= data.matrix(race_survival[,2:101])
y= race_survival$Race
y = as.factor(y)
```
```{r}
set.seed(99)
rf.race_survival = caret::train(x,y,'rf', trControl=trainControl(method='cv',number=5))
```
```{r}
race_info = as.data.frame(table(race_survival$Race))
sum(race_info$Freq)
length(race_info$Var1)
summary(race_info$Freq)
```
```{r}
race_confmat = confusionMatrix(rf.race_survival, norm="none")
race_proptest = propTest(race_confmat$table)
race_confmat
race_proptest
```

#Tumour Mutational Burden
####Survival cases
```{r}
#Model NMF representation by TMB 
tmb_factors_survival = survival_combo
tmb_factors_survival$tmb_cutoff <- cut(tmb_factors_survival$Tumour_Mutational_Burden, c(-Inf,28.0,142.0,Inf), c("Low", "Med", "High"))
tmb_factors_survival = tmb_factors_survival[!is.na(tmb_factors_survival$tmb_cutoff),]

#Make our test and train data
x= data.matrix(tmb_factors_survival[,2:101])
y= tmb_factors_survival$tmb_cutoff
y = as.factor(y)
```
```{r}
set.seed(99)
rf.NMF.tmb_survival = caret::train(x,y,'rf',trControl=trainControl(method='cv',number=5))
```
```{r}
tmb_info = as.data.frame(table(tmb_factors_survival$tmb_cutoff))
sum(tmb_info$Freq)
length(tmb_info$Var1)
summary(tmb_info$Freq)
```
```{r}
tmb_confmat = confusionMatrix(rf.NMF.tmb_survival, norm="none")
tmb_proptest = propTest(tmb_confmat$table)
tmb_confmat
tmb_proptest
```


#Compile Data for survival
```{r}
organ.cv = rf.NMF.organ_survival[["resample"]]
organ.cv = dplyr::select(organ.cv, c("Accuracy", "Kappa"))
organ.cv$Model <- 'Organ of Origin'

subtype.cv = rf.NMF.subtype_survival[["resample"]]
subtype.cv = dplyr::select(subtype.cv, c("Accuracy", "Kappa"))
subtype.cv$Model <- 'TCGA Subtype'

stage.cv = rf.stage_survival[["resample"]]
stage.cv = dplyr::select(stage.cv, c("Accuracy", "Kappa"))
stage.cv$Model <- 'Stage'

grade.cv = rf.grade_survival[["resample"]]
grade.cv = dplyr::select(grade.cv, c("Accuracy", "Kappa"))
grade.cv$Model <- 'Grade'

age.cv = rf.age_survival[["resample"]]
age.cv = dplyr::select(age.cv, c("Accuracy", "Kappa"))
age.cv$Model <- 'Age'

sex.cv = rf.sex_survival[["resample"]]
sex.cv = dplyr::select(sex.cv, c("Accuracy", "Kappa"))
sex.cv$Model <- 'Sex'

immune.cv = rf.NMF.immune_survival[["resample"]]
immune.cv = dplyr::select(immune.cv, c("Accuracy", "Kappa"))
immune.cv$Model <- 'Immune Landscape'

tme.cv = rf.TME_all_survival_withstad[["resample"]]
tme.cv = dplyr::select(tme.cv, c("Accuracy", "Kappa"))
tme.cv$Model <- 'Tumour Microenvironment Score'

reccurence.cv = rf.recurrent_survival[["resample"]]
reccurence.cv = dplyr::select(reccurence.cv, c("Accuracy", "Kappa"))
reccurence.cv$Model <- 'Recurrence'

race.cv = rf.race_survival[["resample"]]
race.cv = dplyr::select(race.cv, c("Accuracy", "Kappa"))
race.cv$Model <- 'Race'

tmb.cv = rf.NMF.tmb_survival[["resample"]]
tmb.cv = dplyr::select(tmb.cv, c("Accuracy", "Kappa"))
tmb.cv$Model <- 'TMB'

```
```{r}
#Tumour molecular characteristics
tumour_model_data_survival = rbind(organ.cv, subtype.cv)
tumour_model_data_survival = rbind(tumour_model_data_survival, immune.cv)
tumour_model_data_survival = rbind(tumour_model_data_survival, tme.cv)
tumour_model_data_survival = rbind(tumour_model_data_survival, tmb.cv)

#Clinicopathologic characteristics
patient_model_data_survival = rbind(age.cv,sex.cv) 
patient_model_data_survival = rbind(patient_model_data_survival, race.cv)
patient_model_data_survival = rbind(patient_model_data_survival, stage.cv)
patient_model_data_survival = rbind(patient_model_data_survival, grade.cv)
patient_model_data_survival = rbind(patient_model_data_survival, reccurence.cv)
```


```{r}
#Plot
library(Rmisc)
#Tumour model
tumour_model_data_survival$Model = as.factor(tumour_model_data_survival$Model)

tumour_model_data_accuracy <- summarySE(tumour_model_data_survival, measurevar="Accuracy", groupvars="Model")
tumour_model_data_kappa <- summarySE(tumour_model_data_survival, measurevar="Kappa", groupvars="Model")
colnames(tumour_model_data_accuracy)[which(names(tumour_model_data_accuracy) == "Accuracy")] <- "Value"
colnames(tumour_model_data_kappa)[which(names(tumour_model_data_kappa) == "Kappa")] <- "Value"

tumour_model_data_accuracy$Metric = 'Accuracy' 
tumour_model_data_kappa$Metric = 'Kappa' 

tumour_model_metrics_survival = rbind(tumour_model_data_accuracy,tumour_model_data_kappa)
tumour_model_metrics_survival$Metric = as.factor(tumour_model_metrics_survival$Metric)

levels(tumour_model_metrics_survival[(tumour_model_metrics_survival$Model) == 'Tumour Microenvironment Score',]) <- 'TME Score'

levels(tumour_model_metrics_survival$Model)[levels(tumour_model_metrics_survival$Model)=="Tumour Microenvironment Score"] <- "TME Score"

#Patient model
patient_model_data_survival$Model = as.factor(patient_model_data_survival$Model)

patient_model_data_accuracy <- summarySE(patient_model_data_survival, measurevar="Accuracy", groupvars="Model")
patient_model_data_kappa <- summarySE(patient_model_data_survival, measurevar="Kappa", groupvars="Model")
colnames(patient_model_data_accuracy)[which(names(patient_model_data_accuracy) == "Accuracy")] <- "Value"
colnames(patient_model_data_kappa)[which(names(patient_model_data_kappa) == "Kappa")] <- "Value"

patient_model_data_accuracy$Metric = 'Accuracy' 
patient_model_data_kappa$Metric = 'Kappa' 

patient_model_metrics_survival = rbind(patient_model_data_accuracy,patient_model_data_kappa)
patient_model_metrics_survival$Metric = as.factor(patient_model_metrics_survival$Metric)

patient_model_metrics_survival$Model <- factor(patient_model_metrics_survival$Model,levels = c( "Age", "Race", "Recurrence", "Sex", "Grade", "Stage"))
```


#95% CI data =  might not need
```{r}
tumour_model_metrics_survival_conf = tumour_model_metrics_survival
tumour_model_metrics_survival_conf$upper = tumour_model_metrics_survival_conf$Value + tumour_model_metrics_survival_conf$ci
tumour_model_metrics_survival_conf$lower = tumour_model_metrics_survival_conf$Value - tumour_model_metrics_survival_conf$ci
tumour_model_metrics_survival_conf
patient_model_metrics_survival_conf = patient_model_metrics_survival
patient_model_metrics_survival_conf$upper = patient_model_metrics_survival_conf$Value + patient_model_metrics_survival_conf$ci
patient_model_metrics_survival_conf$lower = patient_model_metrics_survival_conf$Value - patient_model_metrics_survival_conf$ci
patient_model_metrics_survival_conf

```

###FINAL SURVIVAL FACTOR MODEL PLOTS
```{r}
# Error bars represent 95% CI
tumour_gg_survival = ggplot(tumour_model_metrics_survival, aes(x=reorder(Model,-Value), y=Value, fill=Metric)) + 
    geom_bar(position=position_dodge(), stat="identity", colour= "black") +
    geom_errorbar(aes(ymin=Value-ci, ymax=Value+ci),
                  width=.4,                  
                  position=position_dodge(.9)) + 
  scale_fill_manual("Metric", values = c("Accuracy" = "skyblue3", "Kappa" = "seagreen3")) + 
  scale_y_continuous(limit = c(0, 1)) +
  ggtitle("Tumour Characteristics") + 
  ylab("Value") +
  theme_classic()  + 
  theme(axis.text.x = element_text(colour="black", size = 15,angle = 45, hjust=1)) +
  theme(axis.text.y = element_text(colour="black",size = 15)) + 
  theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_text(colour="black", size=15)) +
  theme(legend.position = "none" ) 


ggsave("tumour_gg_survival_final_new.svg", tumour_gg_survival,width=4, height=5)

ggsave("tumour_gg_survival_final_new.png", tumour_gg_survival, width=6, height=5)

patient_gg_survival = ggplot(patient_model_metrics_survival, aes(x = factor(Model, level = c('Age', 'Race', 'Recurrence','Sex','Grade','Stage' )), y=Value, fill=Metric)) + 
    geom_bar(position=position_dodge(), stat="identity", colour= "black") +
    geom_errorbar(aes(ymin=Value-ci, ymax=Value+ci),
                  width=.4,                  
                  position=position_dodge(.9)) +
  scale_fill_manual("Metric", values = c("Accuracy" = "skyblue3", "Kappa" = "seagreen3")) + 
  scale_y_continuous(limit = c(0, 1)) +
  ggtitle("Clinicopathologic Characteristics") +
  ylab("Value") +
  theme_classic()  + 
  theme(axis.text.x = element_text(colour="black", size = 15,angle = 45, hjust=1)) +
  theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_text(colour="black", size=15)) +
  theme(legend.title = element_text(color = "black", size = 15),
        legend.text = element_text(color = "black", size = 15), 
        legend.position = c(0.8,0.87), 
        legend.background = element_rect(size=0.5, linetype="solid", 
                                  colour ="black"))

ggsave("patient_gg_survival_final_new.svg", patient_gg_survival, width=4, height=5)

ggsave("patient_gg_survival_new.png", patient_gg_survival, width=6, height=5)

library(patchwork)

combined_factor_plots = tumour_gg_survival + patient_gg_survival

combined_factor_plots[[1]] = combined_factor_plots[[1]] + 
                              theme(legend.position = "none" ) + 
                              theme(panel.grid.major.y = element_line(colour = "gray", size = 0.5)) +
                              theme(panel.grid.minor.y = element_line(colour = "gray", size = 0.25)) +
                              theme(plot.margin = unit(c(0.2,5.5,5.5,25), "points"))    

combined_factor_plots[[2]] = combined_factor_plots[[2]] + 
                                  theme(axis.text.y = element_blank(),
                                        axis.ticks.y = element_blank(),
                                        axis.title.y = element_blank(),
                                        axis.line.y = element_blank()) + 
                                  theme(panel.grid.major.y = element_line(colour = "gray", size = 0.5)) +
                                  theme(panel.grid.minor.y = element_line(colour = "gray", size = 0.25)) + 
                                  theme(plot.margin = unit(c(0.2,5.5,5.5,5.5), "points"))   
                                  
combined_factor_plots = combined_factor_plots + theme(text=element_text(size=12)) + theme(panel.grid.major.y = element_line(colour = "gray", size = 0.5)) 



ggsave("combined_factor_plots_new.png", combined_factor_plots, width=7, height=6)
ggsave("combined_factor_plots_new.svg", combined_factor_plots, width=7, height=6)

```